---
title: "Housing in London"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup environment

* Clear all variables in workspace

```{r }
rm(list = ls()) 
```
* Install packages

Packages used: *tidyverse, janitor, readr, dplyr, skimr, DataExplorer,lubridate, ggplot2, geofacet, stringr, fpp2, viridis, patchwork, gghighlight, PerformanceAnalytics*

```{r message=FALSE, warning=FALSE, include=FALSE}
# Set up environment
library(tidyverse)
library(janitor)
library(readr)
library(dplyr)
library(skimr)
library(DataExplorer)
library(lubridate)
library(ggplot2)
library(geofacet)
library(gganimate)
library(ggthemes)
library(stringr)
library(fpp2)
library(patchwork)
library(viridis)
library(gghighlight)
library(PerformanceAnalytics)

```

## Import Data

Housing in London data sets can be found here: [https://www.kaggle.com/justinas/housing-in-london](https://www.kaggle.com/justinas/housing-in-london).

```{r include=FALSE}
# Loading csv file
ldh_m <- read_csv("C:/Users/suong/OneDrive/Desktop/LH data/housing_in_london_monthly_variables.csv")
head(ldh_m)
```

## Data Exploration

The data sets contain monthly and yearly data which centered around the housing market of London. In this notebook, I will be looking at the monthly data.

```{r include=FALSE}
colnames(ldh_m)
```
List of variables in a dataset: 

Field name|Description
------|------
**Monthly data**|
*average_price* |Monthly average house prices
*no_of_crimes* | Monthly number of crimes committed
*houses_sold*  | Monthly number of houses sold
*code*/*date*/*area*/*borough_flag* | 


<details>
  <summary>Click for details of **Variables Statistics\textcolor{blue}{blue}**</summary>

Count unique codes for each data frame
```{r}
n_unique(ldh_m$code)
```
```{r}
min(ldh_m$date)
max(ldh_m$date)
```

ldh_m has data available for 1995-01-01 and 2020-01-01 with 45 unique codes.
The year 2020 has incomplete data (only for 01 day) and will be excluded.

```{r}
# Find unique areas
unique(ldh_m$area)

```

```{r}
# Count observations in each area
data.frame(table(ldh_m$area))
```
There is a discrepancy in the number of observations in some areas.
```{r}
table(ldh_m$borough_flag)
```

```{r }
# Find and remove duplicates (if any)
sum(duplicated(ldh_m))
```
There are no duplicates found.

```{r}
arrange(unique(ldh_m[c("code", "area")]))
```
There are few codes and areas that are mismatched.
Areas: "hackney", "enfield", "south east", "tower hamlets" and codes: E12000008, E09000012, E09000010 E09000030.

```{r include=FALSE}
ldh_m %>% 
  filter(area=="enfield") %>% 
  arrange(desc(code, date)) # replicate with wrong code: delete entire row
ldh_m %>% 
  filter(area=="hackney") %>% 
  arrange(desc(code, date)) # replicate with wrong code: delete entire row
ldh_m %>% 
  filter(area=="south east") %>% 
  arrange(code, date) # replicate with wrong code: delete entire row
ldh_m %>% 
  filter(area=="tower hamlets") %>% 
  arrange(code, date)  # replicate with wrong code: delete entire row
```
Let's have a look at the structures of the data we have, using `str(df)` function.
Mean salary is in character type needs to be converted into numeric.

```{r}
str(ldh_m)
```

```{r}
introduce(ldh_m)
```
</details>
<br>

Plot data description
```{r}
## View missing value distribution 
plot_missing(ldh_m)
```

## Data Cleaning and Transformation

**Summary of Data Cleaning:** 

>
* Excluded some variables from the dataset:
+ excluded column "borough_flag": using "code" column instead.
+ excluded data on date 2020-01-01: as this day seems to have incomplete data.  
* Removed row dated 1998-04-01 for "south east" (replicate with wrong code assigned "E09000012")
* Removed row dated 1998-04-01 for	"hackney" (replicate with wrong code	assigned "E12000008")
* Removed row dated  1996-02-01 for	"enfield" (replicate with wrong code	assigned "E09000030")
* Removed row dated 1996-02-01 for	"tower hamlets" (replicate with wrong code assigned "E09000010")
* Filled up houses_sold with median values in replacing to missing values.

```{r data cleaning }
# Remove replicates with wrong codes
m_new <- ldh_m[!(ldh_m$date == "1998-04-01" & ldh_m$area =="south east" & ldh_m$code == "E09000012") & !(ldh_m$date == "1998-04-01" & ldh_m$area =="hackney" & ldh_m$code == "E12000008") & !(ldh_m$date == "1996-02-01" & ldh_m$area =="enfield" & ldh_m$code == "E09000030") & !(ldh_m$date == "1996-02-01" & ldh_m$area =="tower hamlets" & ldh_m$code == "E09000010"),]

# Exclude data date 2020-01-01 as the year is incomplete
m_new <- m_new %>% filter(date != "2020-01-01") %>% select(-"borough_flag")

```
Houses_sold variable has a fairly small amount of missing values, imputation method will be used to replace missing values with mean values.
```{r}
# Filling missing values (imputation method)
hsold_median = median(ldh_m$houses_sold, na.rm=TRUE)
ldh_m$houses_sold[is.na(ldh_m$houses_sold)] = hsold_median
```

```{r}
# Find unique codes
unique(m_new$code)

```

According to ONS [http://statistics.data.gov.uk/atlas/resource](http://statistics.data.gov.uk/atlas/resource) are codes are defined as follow:

**ONS** | **Count from dataset**
--------|-------------------
E09: 33 London Boroughs | 33
E12: 9 Regions | 9 
E13: 2 zones - Inner and Outer London | 2
E92: Country (England) | 1

The data is available at borough, inner/outer, regions,and country levels. 

```{r Count number of codes per each area group}
# Count how many boroughs are available, code starts with "E09"
boroughs <- m_new %>% filter(str_detect(m_new$code, "E09")) 
n_unique(boroughs$code)
unique(boroughs$code)

# Count how many regions are available, code starts with "E12"
regions <- m_new %>% filter(str_detect(m_new$code, "E12")) 
n_unique(regions$code)
unique(regions$code)

# Check if there are inner and outer London data available, code starts with "E13"
greater_ld <- m_new %>% filter(str_detect(m_new$code, "E13")) 
n_unique(greater_ld$code)
unique(greater_ld$code)

```

```{r}
# Add a new column defining inner / outer boroughs
zoneld <- m_new %>% 
  mutate(zone = case_when(
    area %in% c("camden", "greenwich", "hackney", "hammersmith and fulham", "islington",
"kensington and chelsea", "lambeth", "lewisham", "southwark", "tower hamlets", "wandsworth", "westminster") ~ "inner", 
area %in% c("barking and dagenham", "barnet", "bexley","brent","bromley", "croydon", "ealing", "enfield", "haringey", "harrow", "havering", "hillingdon", "hounslow", "kingston upon thames", "merton", "newham", "redbridge", "richmond upon thames","sutton", "waltham forest") ~ "outer",
TRUE ~ "other"))

```

```{r}
# Add new column with extract year value from date 
zoneld <-  zoneld %>% 
mutate(date = ymd(date), year=as.numeric(substr(date, 1, 4)))
```

## Exploratory Data Analysis

Plot histogram of houses price over years
```{r}
ggplot(m_new,aes(x=average_price))+ 
  geom_histogram(colour = "darkgray", fill = "white", bins = 40)+
 labs(xlab = "Steps", ylab = "Count", title = "Histogram of Houses Price Over Years")+
 geom_vline(xintercept = mean(m_new$average_price, na.rm = TRUE), 
          color = "red", show.legend = TRUE, size = 1.5)+
 geom_vline(xintercept = median(m_new$average_price, na.rm = TRUE), 
           color = "blue", show.legend = TRUE, size = 1.5)+
 scale_color_manual(name = "statistics", values = c(mean = "red", median = "blue"))+
scale_x_continuous(labels = scales::comma)
```
```{r fig.height=4, fig.width=8}
# Correlation among variables
corr <- m_new[, -c(1,2,4)]
chart.Correlation(corr, histogram=TRUE, pch=19)

```

```{r}
# Plot inner vs outer London 
m_new <- m_new %>% 
mutate(date = ymd(date),
        year=as.numeric(substr(date, 1, 4)))
m_new %>% 
  filter(str_detect(code, c("E13", "E92"))) %>% 
  group_by(area, date) %>% 
  summarise(mean_price = mean(average_price), .groups = "drop") %>% 
ggplot(aes(x=date, y=mean_price, group=area, colour=area)) +
  geom_line(lwd=.8)+
theme_set(theme_minimal())+
  labs(x= NULL, y="Average price (£)", title = "Average Houses Price London Vs England")+
  scale_y_continuous(labels = scales::comma)+
   scale_x_date(date_breaks = "2 years", date_labels = "%Y")

```
```{r fig.width=10, fig.show="hide"}
# Plot number of houses sold across regions
p8 <- m_new %>% 
  filter(str_detect(code, "E12")) %>% 
  group_by(area, date) %>% 
  summarise(sum_sold = sum(houses_sold, na.rm = TRUE), .groups = "drop") %>% 
ggplot(aes(x=reorder(area, sum_sold), y=sum_sold, group=area, fill=area)) +
  geom_boxplot(show.legend = FALSE)+
theme_set(theme_minimal())+
  labs(x=NULL, y="number of houses sold", title = "Number of Houses Sold Across Regions")+
  scale_y_continuous(labels = scales::comma)
```


```{r fig.width=10, fig.show="hide"}
# Plot average houses price across regions
p12 <- m_new %>% 
  filter(str_detect(code, "E12")) %>% 
  group_by(area, date) %>% 
  summarise(mean_price = mean(average_price), .groups = "drop") %>% 
ggplot(aes(x=reorder(area, mean_price), y=mean_price, group=area, fill=area)) +
  geom_boxplot(show.legend = FALSE)+
theme_set(theme_minimal())+
  labs(x=NULL, y="Average price", title = "Average Houses Price Across Regions")+
  scale_y_continuous(labels = scales::comma)
```

```{r fig.height=6, fig.width=10}
(p8 / p12) + plot_layout(ncol=1)
```

```{r}
change <- m_new %>%
  filter(str_detect(code, "E12")) %>% 
  group_by(area, year) %>% 
  arrange(year, .by_group = TRUE) %>%
  summarise(pct_change = (houses_sold/lag(houses_sold) - 1) * 100, .groups = "drop") %>% 
  ggplot(aes(area, pct_change, fill=area))+
  geom_boxplot()+
  theme(legend.position = "bottom")

change
```

```{r fig.height=5, fig.width=6}
# Plot compare average houses price across London boroughs
p6 <- zoneld %>% 
  filter(str_detect(code, "E09") & zone!="other") %>% 
  group_by(area, date, zone) %>% 
  summarise(mean_price = mean(average_price), .groups = "drop") %>% 
ggplot(aes(x=date, y=mean_price, group=area, color=area)) +
  geom_line(lwd = 0.8)+
theme_set(theme_minimal())+
  labs(x="Date", y="Average price", title = "Average House Price London Boroughs")+
  scale_y_continuous(labels = scales::comma)+
   scale_x_date(date_breaks = "5 years", date_labels = "%Y")+
  theme(legend.position = "bottom")+
 gghighlight(max(mean_price),5, label_key = area)+
  facet_wrap(~zone, ncol=1)
```


```{r fig.height=5, fig.width=6}
# Plot compare number of houses sold London boroughs
p7 <- zoneld %>%
  filter(str_detect(code, "E09") & zone!="other") %>% 
  group_by(area, date, zone) %>% 
  summarise(sum_sold = sum(houses_sold, na.rm = TRUE), .groups = "drop") %>% 
ggplot(aes(x=date, y=sum_sold, group=area, color=area)) +
  geom_line(lwd = 0.8)+
theme_set(theme_minimal())+
  labs(x=NULL, y="number of houses", title = "Number of Houses Sold Across London Boroughs")+
  scale_y_continuous(labels = scales::comma)+
   scale_x_date(date_breaks = "5 years", date_labels = "%Y")+
  theme(legend.position = "bottom")+
 gghighlight(max(sum_sold),5, label_key = area)+
  facet_wrap(~zone, ncol=1)
```

```{r fig.height=6, fig.width=12}
(p6 | p7) + plot_layout(nrow=1)
```

```{r warning=FALSE, message=FALSE}
# Plot average house price London and other regions
p4 <- m_new %>% 
  filter(str_detect(code, c("E92", "E12"))) %>%
  group_by(area, date) %>% 
  summarise(mean_price = mean(average_price), .groups = "drop") %>% 
ggplot(aes(x=date, y=mean_price, group=area, colour=area)) +
  geom_line(lwd = .8)+
theme_set(theme_minimal())+
  labs(x= NULL, y= NULL, title = "Average House Price across regions")+
  scale_y_continuous(labels = scales::comma)+
   scale_x_date(date_breaks = "5 years", date_labels = "%Y")+
  gghighlight(area==c("england", "london", "south east"), label_key = area)
```


```{r}
# Plot compare Average houses price in London and other regions (2)
p5 <- m_new %>% 
  filter(str_detect(code, c("E13", "E12"))) %>% 
  group_by(area, date) %>% 
  summarise(mean_price = mean(average_price), .groups = "drop") %>% 
ggplot(aes(x=date, y=mean_price, group=area, color=area)) +
  geom_line(show.legend = FALSE, lwd=.8)+
  labs(x=NULL, y="Average price", title = "Average Houses Price London vs Other Regions")+
  scale_y_continuous(labels = scales::comma)+
  scale_x_date(date_breaks = "6 years", date_labels = "%Y")+
   theme(legend.position = "bottom")+
  facet_wrap(~area)+
  expand_limits(y = 0)

```

```{r fig.height=6, fig.width=12}
(p4 | p5) + plot_layout(ncol=2, widths =  c(.4, .6))
```
The heatmap chart below shows the *average houses price in each London borough by year*.

```{r fig.height=5.5, fig.width=6}
# Tile map of Average House Price by London boroughs
p1 <- zoneld%>%
  filter(str_detect(code, "E09")) %>%
  group_by(area, zone, year) %>% 
  summarise(mean_price = mean(average_price)/1000, .groups = "drop") %>% 
  ggplot(mapping=aes(x = year, 
                     y = factor(area, levels = rev(levels(factor(area)))))) +
  geom_tile(aes(fill = mean_price), colour = "white", height = 1, width = 1) +
  #labs(title = "Average Houses Price by London Boroughs",
      # x = NULL, y = NULL, subtitle = "1995 to 2019") +
  labs(title = "Average Houses Price by London Boroughs",
       x = NULL, y = NULL) +
  scale_fill_viridis_c(name="Avg Price (£ thousands)", 
                     guide=guide_colorbar(frame.colour="black", ticks.colour="black"))
```


```{r fig.height=2, fig.width=5}
# Tile map of Average House Price in Inner / Outer London 
p2 <- m_new%>%
  filter(str_detect(code, "E13")) %>%
  group_by(area, year) %>% 
  summarise(mean_price = mean(average_price)/1000, .groups = "drop") %>% 
  ggplot(aes(x = year, 
                     y = factor(area, levels = rev(levels(factor(area)))))) +
  geom_tile(aes(fill = mean_price), colour = "white", height = 1, width = 1) +
  #labs(title = "Average Houses Price in Inner/Outer London",
     #  x = NULL, y = NULL, subtitle = "1995 to 2019") +
  labs(x = NULL, y = NULL)+
  scale_fill_viridis_c(option = "plasma", name="Avg Price (£ thousands)", 
                     guide=guide_colorbar(frame.colour="black", ticks.colour="black"))+
  theme(legend.position = "top")

```

```{r}
# Plot Average houses price inner vs outer London 
p3 <- m_new %>% 
  filter(str_detect(code, "E13")) %>% 
  group_by(area, date) %>% 
  summarise(mean_price = mean(average_price), .groups = "drop") %>% 
ggplot(aes(x=date, y=mean_price, group=area, colour=area)) +
  geom_line(lwd = .8, show.legend = FALSE)+
 # labs(x= NULL, y= NULL, title = "Average Houses Price Inner vs Outer London")+
  theme_set(theme_minimal())+
  labs(x= NULL, y= NULL)+
  scale_y_continuous(labels = scales::comma)+
   scale_x_date(date_breaks = "5 years", date_labels = "%Y")+
 # theme(legend.position = "bottom")+
  facet_wrap(~area)

```


```{r fig.height=4, fig.width=5}
# Plot average houses price across Inner/Outer boroughs
inner.p <- zoneld %>% 
  filter(str_detect(code, "E09") & zone=="inner") %>% 
  group_by(area, date) %>% 
  summarise(mean_price = mean(average_price, na.rm = TRUE),.groups = "drop") %>% 
  ggplot(aes(x=reorder(area, mean_price), y=mean_price)) +
   geom_bar(stat = "summary", fun = "mean", fill="darkorange2")+
  coord_flip()+
  labs(x=NULL, y=NULL, title = "Average House Price Across Inner Boroughs")+
  scale_y_continuous(labels = scales::comma)
outer.p <-zoneld %>% 
  filter(str_detect(code, "E09") & zone=="outer") %>% 
  group_by(area, date) %>% 
  summarise(mean_price = mean(average_price, na.rm = TRUE),.groups = "drop") %>% 
 ggplot(aes(x=reorder(area, mean_price), y=mean_price)) +
   geom_bar(stat = "summary", fun = "mean", fill="darkslategrey")+
  coord_flip()+
  labs(x=NULL, y=NULL, title = "Average House Price Across Outer Boroughs")+
 scale_y_continuous(labels = scales::comma)

```

```{r fig.height=5, fig.width=12, fig.show="hide"}
# Plot number of houses across inner/outer boroughs
(inner.p | outer.p) + plot_layout(ncol = 2, widths = c(0.6, 0.4))
```

```{r fig.height=10, fig.width=10}
library(patchwork)
p1 / p3/ (inner.p | outer.p)/ p2 + plot_layout(nrow=4, heights = c(0.45, .2, 0.3, 0.05))

```

```{r}
# Tile map of number of houses sold by London boroughs
p9 <- zoneld%>%
  filter(str_detect(code, "E09")) %>%
  group_by(area, zone, year) %>% 
  summarise(sum_sold = sum(houses_sold), .groups = "drop") %>% 
  ggplot(mapping=aes(x = year, 
                     y = factor(area, levels = rev(levels(factor(area)))))) +
  geom_tile(aes(fill = sum_sold), colour = "white", height = 1, width = 1) +
 # labs(title = "Number of Houses Sold by London Boroughs",
      # x = NULL, y = NULL, subtitle = "199./9. to 2018/19") +
  labs(title = "Number of Houses Sold by London Boroughs",
       x = NULL, y = NULL) +
 # theme(legend.position = "top")+
  scale_fill_viridis_c(option = "viridis", name="Number of houses",
                     guide=guide_colorbar(frame.colour="black", ticks.colour="black"))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Plot compare number houses sold in Inner/Outer London
p10 <- m_new %>% 
  filter(str_detect(code, "E13")) %>% 
  group_by(area, date) %>% 
  summarise(sum_sold = sum(houses_sold), .groups = "drop") %>% 
ggplot(aes(x=date, y=sum_sold, group=area, color=area)) +
  geom_line(show.legend = FALSE, lwd=.8)+
theme_set(theme_minimal())+
  #labs(x=NULL, y="Number of houses", title = "Number Houses Sold Inner Vs Outer London")+
   labs(x=NULL, y=NULL)+
  scale_y_continuous(labels = scales::comma)+
  scale_x_date(date_breaks = "3 years", date_labels = "%Y")+
   theme(legend.position = "bottom")+
  facet_wrap(~area)+
  expand_limits(y = 0)
p10
```


```{r fig.height=2, fig.width=6}
# Tile map of number of houses sold in Inner and Outer London
p11 <- m_new%>%
  filter(str_detect(code, "E13")) %>%
  group_by(area, year) %>% 
  summarise(sum_sold = sum(houses_sold), .groups = "drop") %>% 
  ggplot(mapping=aes(x = year, 
                     y = factor(area, levels = rev(levels(factor(area)))))) +
  geom_tile(aes(fill = sum_sold), colour = "white", height = 1, width = 1) +
  #labs(title = "Number of Houses Sold in Inner/Outer London",
  #  x = NULL, y = NULL, subtitle = "1995 to 2019") +
  labs(x = NULL, y = NULL) +
  theme(legend.position = "bottom")+
  scale_fill_viridis_c(option = "viridis", name="Number of Houses", 
                     guide=guide_colorbar(frame.colour="black", ticks.colour="black"))
```

```{r boxplot, fig.height=4, fig.width=5}
# Plot number of houses sold across Inner/Outer boroughs
inner <- zoneld %>% 
  filter(str_detect(code, "E09") & zone=="inner") %>% 
  group_by(area, year, zone) %>% 
  summarise(sum_sold = sum(houses_sold, na.rm = TRUE),.groups = "drop") %>% 
  ggplot(aes(x=reorder(area, sum_sold), y=sum_sold)) +
  geom_bar(stat = "summary", fun = "mean", fill="orange")+
  coord_flip()+
  labs(x=NULL, y=NULL, title = "Houses Sold Across Inner Boroughs")+
  scale_y_continuous(labels = scales::comma)
outer <-zoneld %>% 
  filter(str_detect(code, "E09") & zone=="outer") %>% 
  group_by(area, year, zone) %>% 
  summarise(sum_sold = sum(houses_sold, na.rm = TRUE),.groups = "drop") %>% 
 ggplot(aes(x=reorder(area, sum_sold), y=sum_sold)) +
  geom_bar(stat = "summary", fun = "mean", fill="darkslategrey")+
  coord_flip()+
  labs(x=NULL, y=NULL, title = "Houses Sold Across Outer Boroughs")+
 scale_y_continuous(labels = scales::comma)

```

```{r fig.height=5, fig.width=12, fig.show="hide"}
# Plot number of houses across inner/outer boroughs
(inner | outer) + plot_layout(ncol = 2, heights = c(0.4, 0.6))
```

Put all number of houses sold in one graph.
```{r fig.height=10, fig.width=12, warning=FALSE, message=FALSE}
p9 / p10 / (inner | outer) / p11 + plot_layout(nrow=4, heights = c(0.45, .2, 0.3, 0.05))

```

### Preliminary Analysis

The data covers from month 1:1995 to month 1:2019, resulting in 299 observations.
This time series analysis is using these 299 observations.

```{r}
# Filtering average house price values for London
ld <- m_new %>% 
 filter(str_detect(code, "E09")) %>% 
  mutate(average_price = average_price/1000)

# Creating a time series of average house price
ldmonth.ts <- ts(ld[, 3], start = c(1995,1), end = c(2019, 12), frequency = 12)

# Plot
ts1 <- autoplot(ldmonth.ts, lwd = 0.8, col = "blue")+
  ggtitle("Time plot: Average London House Price Per Month")+
  ylab("Thousands of 2019  GBP Pounds")

```
Investigate transformation
```{r}
# Take the first difference of the data
Dldmonth.ts <- diff(ldmonth.ts)

# Plot
ts2 <- autoplot(Dldmonth.ts)+
  ggtitle("Time plot: Change in Average London House Price Per Month")+
  ylab("Thousands of 2019  GBP Pounds")+
  xlab("1:1995 to 12:2019")

```

```{r}
# Series appears trend-stationary, use to investigate seasonality
ts3 <- ggseasonplot(Dldmonth.ts)+
  ggtitle("Change in Monthly Average House Price")+
  ylab("Thousands of 2019 GBP Pounds")+
  theme(legend.position = "bottom")

```

```{r}
# Let's look at another seasonal plot, the subseries plot
ts4 <- ggsubseriesplot(Dldmonth.ts)+
  ggtitle("")+
  ylab("Thousands of 2019 GBP Pounds")
```

```{r fig.height=10, fig.width=12}
(ts1 + ts2) / (ts3+ts4)
```

**Observation:**
* Our series, ldmonth.ts, has (or not has) trend and seasonality. => coi ky lai roi rut ra kl
* To remove the trend, we take the first difference.
* The first differenced series still has seasonality.

```{r fig.height=7, fig.width=7}
# Filter London values
ld1 <- m_new %>% filter(area=="london")
#unique(ld1$code)
# Creating a time series
ld1month.ts <- ts(ld1[, 3], start = c(1995,1), end = c(2019, 12), frequency = 12)
# Decompose
london_decomposed <- decompose(ld1month.ts)
plot(london_decomposed, col="blue", lwd = .8)

```

The average house price of Greater London between 1995 and 2019 and inferences around the overall trend and seasonality.

## Visualization

```{r}
#create London Boroughs grid
london_grid <- tribble(~code, ~row, ~col, ~name,
                       "E09000001",	4,	5,	"City of London",
                       "E09000002",	4,	8,	"Barking and Dagenham",
                       "E09000003",	2,	4,	"Barnet",
                       "E09000004",	5,	8,	"Bexley",
                       "E09000005",	3,	3,	"Brent",
                       "E09000006",	6,	6,	"Bromley",
                       "E09000007",	3,	4,	"Camden",
                       "E09000008",	6,	5,	"Croydon",
                       "E09000009",	3,	2,	"Ealing",
                       "E09000010",	1,	5,	"Enfield",
                       "E09000011",	5,	7,	"Greenwich",
                       "E09000012",	3,	6,	"Hackney",
                       "E09000013",	4,	2,	"Hammersmith and Fulham",
                       "E09000014",	2,	5,	"Haringey",
                       "E09000015",	2,	3,	"Harrow",
                       "E09000016",	3,	8,	"Havering",
                       "E09000017",	3,	1,	"Hillingdon",
                       "E09000018",	4,	1,	"Hounslow",
                       "E09000019",	3,	5,	"Islington",
                       "E09000020",	4,	3,	"Kensington and Chelsea",
                       "E09000021",	5,	2,	"Kingston upon Thames",
                       "E09000022",	5,	4,	"Lambeth",
                       "E09000023",	5,	6,	"Lewisham",
                       "E09000024",	6,	4,	"Merton",
                       "E09000025",	4,	7,	"Newham",
                       "E09000026",	3,	7,	"Redbridge",
                       "E09000027",	6,	3,	"Richmond upon Thames",
                       "E09000028",	5,	5,	"Southwark",
                       "E09000029",	7,	4,	"Sutton",
                       "E09000030",	4,	6,	"Tower Hamlets",
                       "E09000031",	2,	6,	"Waltham Forest",
                       "E09000032",	5,	3,	"Wandsworth",
                       "E09000033",	4,	4,	"Westminster"
)

```

```{r}
#merge
ld_houses <- left_join(m_new, london_grid, by = c("code" = "code"))
ld_houses <- ld_houses %>% 
mutate(date = ymd(date),
  year=as.numeric(substr(date, 1, 4))) 
ld_houses
```

```{r fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
# plot
ld_houses %>% 
 filter(str_detect(code, "E09")) %>% 
  group_by(name, year) %>% 
  summarise(mean_price = mean(average_price), .groups = "drop") %>%
  ggplot(aes(x=year, y=mean_price)) +
geom_col(fill="darkcyan")+
  scale_x_continuous(breaks = c(1996, 2016), labels = c(1996, 2016)) +
  scale_y_continuous(breaks = c(500000, 1000000, 1500000), labels = c("0.5M","1M","1.5M")) +
facet_geo(~name, grid = london_grid)+
  labs(title = "Housing Price trends in London Boroughs 1995 - 2019",
       caption = "Annualised average price",
       x = "1995 - 2019",
       y = "Annualised price") +
 theme(strip.text.x = element_text(size = 8.5, margin=margin(0.1,0,0.1,0,"cm")),
       strip.background = element_rect(fill = "slategray2"),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank())

```

```{r fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
# plot number of houses sold
ld_houses %>% 
 filter(str_detect(code, "E09")) %>% 
  group_by(name, year) %>% 
  summarise(sum_sold = sum(houses_sold, na.omit = TRUE), .groups = "drop") %>% 
  ggplot(aes(x=year, y=sum_sold)) +
geom_col(fill="deepskyblue4")+
  scale_x_continuous(breaks = c(1996, 2016), labels = c(1996, 2016)) +
  scale_y_continuous(breaks = c(500000, 1000000, 1500000), labels = c("0.5M","1M","1.5M")) +
facet_geo(~name, grid = london_grid)+
  labs(title = "Houses Sold trends in London Boroughs 1995 - 2019",
       caption = "Annualised average price",
       x = "1995 - 2019",
       y = "Annualised price") +
 theme(strip.text.x = element_text(size = 8.5, margin=margin(0.1,0,0.1,0,"cm")),
       strip.background = element_rect(fill = "slategray2"),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank())

```
